Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-08-01 Document/paper.tex
edits RP

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{subfig}
\usepackage{array}

\usepackage{multirow}
\usepackage{multicol}
\usepackage{blindtext}
\newcolumntype{?}{!{\vrule width 1pt}}
\newcommand{\Msun}{\,{\rm Mpc}$_{\odot}$\,}
\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
\newcommand{\kpch}{\,{\rm kpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
\newcommand{\kpc}{\,{\rm kpc}\,}

\title[Dark matter halo shapes at $z=0$] {Dark matter halo shapes at
  $z=0$ in the Auriga simulations} 
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'isica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^{2}$Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzschild-Str. 1, D-85741 Garching, Germany\\
}

@@

% Abstract of the paper
\begin{abstract}
We present shape measurements of Milky Way sized dark matter halos at
redshift $z=0$ in a suite of 30 zoom simulations from the Auriga
project. 
We compare the results in full magnetohydrodynamics against dark
matter only simulations. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
At distances $\lesssim 30$ kpc, rounder dark matter distributions
correlate with extended massive stellar discs and low core gas 
densities.  
We measure the alignment between the halo and the stellar disc at
different radii and find almost perfect alignment at
$0.25R_{200}\sim56$ kpc.    
In some cases the alignment significantly changes as a function of
radius implying that the halo shape twists; 
this effect correlates with extended stellar discs and
low gas densities and is almost absent in the dark matter only
simulations.  
In a comparison against observational constraints we find that $20\%$
of halos in our sample are consistent with observational results derived
from the Pal 5 stream that favor an almost spherical shape.
Including baryons is a required element to achieve this
level of agreement. In contrast, none of the simulations (dark matter
only nor with baryons) match the constrains derived from the
Sagittarius stream favoring an oblate dark matter halo.
\end{abstract}

@@

Our physical picture of the Universe as a whole has be shaped by
accurate observations and modeling of our own Galaxy. 
Explaining the matter budget and kinematical state of the Milky Way
(MW) is  equivalent to figuring out its formation process in a
cosmological context. 
As a first approximation, the MW morphology and
separation into global kinematically coherent components, such as its
disc and bulge, can be used to support the  existence of a Dark Matter
(DM) component to explain its dynamics around the solar neighborhood
\citep{2000MNRAS.311..361O,2009PASJ...61..227S,2010JCAP...08..004C,2013ApJ...779..115B,Iocco15}.

A more detailed description of the full three-dimensional MW
gravitational potential is possible through the interpretation of 
fossil records of stellar streams.
These remanents, resulting from infalling globular clusters or
satellite galaxies that got tidaly disrutpted by the gravitational
potential of the Milky Way, allow in principle tighter constrains 
on the shape of the dark matter halo in outter regions of our Galaxy
\citep{1998ApJ...495..297J,1999MNRAS.307..495H, 1999MNRAS.307..877T}
than dynamical measurements close to the solar neighborhodd.

The observational constraints on the gravitational potential shape can then
be confronted against the expectations from different galaxy formation
models in an explicit cosmological context.
For instance, in the current dominant paradigm of Cold Dark Matter
(CDM) dominated Universe galaxies are expected to be hosted by
triaxial DM halos. To what extent the CDM expectations are born out by
observations in our Galaxy? Can the MW's DM halo shape be considered
typical or atypical in a cosmological context?

These two questions have been challening to address because for a long
time it was difficult to produce realistic galactic discs within the
CDM context \citep{1997ApJ...478...13N}.
The success of such enterprise has been the result of high numerical
resolution and the understanding that baryonic effects such as stellar
feedback and black hole feedback play an important role in forming a
galactic disc resembling the Milky Way.
Along the way, many numerical experiments have also show that the baryonic
effects also impact the dark matter halo shape making the halo rounder
than it otherwise be in a dark matter only simulation
\citep{Dubinski94,Debattista08,Kazantzidis10,Abadi10,Bryan13,Chua19,Artale19}. 
But only few simulations have presented detailed results on another
important aspect of the disc-halo relationship, namely, the radial
evolution of the alignment between the stellar disc and the dark
matter halo \citep{Bailin05,Debattista13}.

@@

In this paper we analyze the internal DM halo shape in thirty MW type
galaxies from state-of-the-art simulations from the Auriga project.
These simulations have large enough numerical resolution, and explicit
cosmological context and an appropriate feedback physics to produce
realistic MW discs.
Futhermore, the relatively large number of simulated systems allows us
to have a handle on the statistical significance of our results.
This paper is structured as follows. 
In Section \ref{sec:numerical} we present the most relevant details of
the simulations, in Section \ref{sec:method} we present the method
that we use to measure the DM halo shape. 
In Section \ref{sec:results} we present our results focusing on the
radial shape trends at $z=0$ and the alignments with the stellar disc.
We discuss our results in Section \ref{sec:discussion} to place them
into the context of other numerical work, explore correlations of the
shape with baryonic properties in the disc and finally make a direct
comparison against observational constraints for the MW's dark matter halo shape.
We finalize with our conclusions in Section \ref{sec:conclusions}.

@@

The 30 zoom-in halos have a dark matter particle mass of $\sim 3\times
10^5$\Msun while the barynic mass resolution is $\sim 5\times 10^4$\Msun.
The softening lenght for gravitational force computation for stellar
particles and high-resolution dark matter particles 
is fixed to be 500 $h^{-1}$ pc in comoving coordinates up to $z=1$,
and 396 pc in physical coordinates afterwards.
The gravitational softening lenght for gas cells changes with the mean
cell radius but is limited to be larger than the stellar softening
lenght and 1850 pc physical. 
From now on we refer to the haloes simulated with baryonic physics as the
MHD sample and to the haloes simulated with dark matter only as the
DMO sample.

\begin{figure*}
  \centering
  \subfloat[DMO simulation. Shape at small
    radius.]{\includegraphics[width=0.5\textwidth]{level4_DM_halo_24_2.png}}  
  \hfill
  \subfloat[DMO simulation. Shape at large
    radius.]{\includegraphics[width=0.5\textwidth]{level4_DM_halo_24_4.png}}  
  \hfill

  \subfloat[MHD simulation. Shape at small
    radius.]{\includegraphics[width=0.5\textwidth]{level4_MHD_halo_24_2.png}}  
  \hfill
  \subfloat[MHD simulation. Shape at large
    radius.]{\includegraphics[width=0.5\textwidth]{level4_MHD_halo_24_4.png}}  
  \hfill 
  \caption{DM density in logarithmic scale within a slice of one tenth
    of the virial radius.
    The cut is perpendicular to the short axis of the inertia tensor ellipsoid.
    The black ellipses show the results of the fitting procedure
    described in Section \ref{sec:method}. 
    Upper/lower panels correspond to DMO/MHD simulations, respectively.
    Left/right panels show data at small/large radii, respectively.
    This plot showcases the most noticeable effect in all halos
    across the Auriga simulations: haloes are rounder at all radii
    after baryonic physics is included.}
\label{fig:slices}
\end{figure*}

\begin{figure*}
\begin{center}
\subfloat[DMO simulations.]{\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}}
\subfloat[MHD simulations.]{\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}}
\end{center}
\caption{Axial ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $14\pm 1$ kpc
  ($230\pm 15$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder than its DMO counterparts.
  Third, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend).}
  \label{fig:triaxiality_plane}
\end{figure*}

\begin{figure*}
\subfloat[DMO simulations.]{\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_DM.pdf}}
\subfloat[MHD simulations.]{\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_MHD.pdf}}
\caption{Triaxility cumulative distribution at five different radii.
  Right/left panel corresponds to DMO/MHD simulations, respectively. 
  In DMO simulations the median triaxiality at all radii is larger
  than $2/3$. Furthermore, the triaxility increases as one moves
  towards the inner part of the halo.
  In MHD simulations these trend reverses.
  The median triaxility at all radii is smaller than $1/3$ and the
  halo is less triaxial as on moves towards the stellar disc.}
\label{fig:triaxial_cumulative}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{angles_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{Angles between the principal axis of the dark matter halo
  shape and the stellar disc shape as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major/middle/minor axis.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value at every radius.
  In the upper row the haloes come from the DMO simulation, 
  showing that the directions in the ellipsoids describing the shape
  are constant as a function of radius for the most part of the sample.
  In the lower row the haloes come from the MHD simulation providing a 
  self-consistent comparison with the stellar discs. 
  In this case the dark matter shells twists significantly in a
  noticeable fraction of the haloes.
  Interestingly, an almost perfect halo-disc alignment happens across
  the sample at an intermediate radius of $0.25R_{200}$ ($56\pm
  4$kpc). 
}
\label{fig:cumulative_alignment}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{correlations_twisting_triaxiality_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alignment (measured as the standard deviation
  of the alignment angles between two minor shape axis)
  as a function of the halo triaxialiy at the radius indicated
  in each panel's title.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient while the line shows the result of the
  minimum squares fit to the data.
  Twisting does not correlate with triaxility in the same way at every
  radii. The strongest correlation appears around $0.12R_{200}$ ($28\pm2$ kpc).} 
\label{fig:alignment_correlations}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the same
  baryonic disc properties in Figure \ref{fig:disc_correlations}.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best minimum squares fit to a line.
  In this case the disc to total mass ratio does not correlate stronly
  with the alignment twisting.
  The gas density at the center and the disc optical radius still show
  a strong correlation.}
\label{fig:alignment_correlations}
\end{figure*}

@@

The DM halo shape is an estimate of either the isopotential or
isodensity surfaces.   
Observational inference models usually estimate the 
isopotential contours which are probed by tracers (gas, stars), while
simulations work with the isodensity contours which can be directly
calculated from particle positions.  
Furthermore, the density contours in thin shells are very sensitive to
the presence of small satelites.  
For this reason we measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.  
This method yields results in good agreement to the isodensity
contours for radii $\leq 140$ kpc as explored by
\citep{VeraCiro11}.

@@

Following the convergence criterion by \cite{VeraCiro11} in DM only simulations
we restrict the measurement of the ellipsoidal parameters to radii
between $2$kpc and $R_{200}$ where  $R_{200}$ corresponds to the radius
computed enclosing a sphere with 200 times the critical density of the
Universe. 
This scale is similar to the gravitational smoothing lenght for
stellar particles. 
For this reason, in order to make the DM halo shape results comparable
against stellar disc shape measurements, we report our DM halo shape
results above $10$ kpc.
For the shape of the stellar disk we include all stellar particles
within a radius of $0.01R_{200}$, that is $\approx 23$kpc.

@@

We summarize this trend in Figure \ref{fig:triaxiality_plane} by
plotting the results of all the 30 halos in the DMO sample.
The left panel shows every halo in the $c/a$-$b/a$ plane at
two different radii $R_{200}/8 (\sim 20$kpc$)$ and $R_{200}$. 
The outer part of the halo is systematically rounder than its inner
region. 
Nevertheless the halo shape can still be considerated to be prolate at
all radii.

@@

The left panel of Figure \ref{fig:triaxial_cumulative} shows that in
the DMO sample the triaxility has a median larger than $2/3$ (a
typical value that marks the transition from triaxiality to
prolateness) at all radii. 
Furthermore this median value increases as we move towards the inner part of the
halo.
The right panel shows how the correlations in the MHD simulations go
in the opposite direction with respect to DMO result.s
There the median triaxility is alwas smaller than $1/3$ (a typical
values that marks the transition from triaxility to prolateness) and this
median value decreases as we move towards the inner part of the halo.

We now quantify this change in triaxility at the level of individual halos.
We compute $\Delta T\equiv T_{\rm MHD}-T_{\rm DMO}$ to quantify the
change between the triaxility in the MHD and the DMO simulation.
Figure \label{fig:delta_triaxial_cumulative} shows the cumulative
distribution at the same radii as in
Figure \label{fig:delta_triaxial_cumulative}.  
From this Figure we observe that at $R_{200}$ all the halos have
reduced its triaxility include baryonic physics.
At smaller radii this general trend continues, although there is a
small fraction ($\approx 1/15$) of halos that slightly increased their
triaxility with MHD physics. 
However, this small fraction correspond to halos that already were
outliers in the DMO sample and had triaxilities around $1/3$,
considerably lower than its parent sample.

@@

A common assumption in dynamical models of the MW DM halo is that
its minor axis is perfectly aligned with the stellar disc minor axis.
Although it is a reasonable assumption to guarantee the stability of
the galactic disc in simplified models of isolated galaxies, this
might not perfectly hold in an explicit cosmological context. 
To examine the degree of validity of this assumption we study in this
subsection the alignment between the eigenvectors of the inertia tensor of
stellar particles within $0.1R_{200}$ ($\sim 23\pm 2$ kpc) and the
eigenvectors of the dark matter halo shape.

In Figure \ref{fig:cumulative_alignment} we summarize our main results
regarding these alignments with the halo shape measured at five
different radii.
The upper row shows the alignment of the halos in the DMO simulations
with the stellar disc in the MHD simulations.
The main objective of this measurement is to calibrate the radial
evolution of the DM halo shape. 
We find that the DM shape remains constant with radius.

The upper row in Figure \ref{fig:cumulative_alignment} shows 
the alignments between the major/intermediate/minor axis of the
disk and the halo as a function of the dimensionless radius at which
the halo shape was measured.
The upper row shows the angle between the DMO halo and the stellar
disk in the MHD simulation.
This comparison help us to measure how the shape directions in the
dark matter shells change with radius.
For the bulk of the sample the lines are horizontal.
Only two halos in the sample show a significant change in alignment in
on the three shape axis.

The lower row in Figure \ref{fig:cumulative_alignment} shows th
results for the MHD simulations.
The most striking feature of this plot is that around the radius of
$0.25R_{200}$ there is an almost perfect halo-disc alignment measured
in three shape axis.
Above and below this radius the degree of alignment changes.
We find the strongest missalignments at a radius of $0.06R_{200}$ when
the median and the standard deviation of the alignment angles between
the minor/median/major axis are $18\pm18$, $25\pm22$ and $7\pm21$
degrees, respectively;  at $0.25R_{200}$ these values drop to $2\pm3$,
$2\pm4$ and $1\pm 3$ degrees, respectively

This changing alignment as a function of radius can be seen as a
twisting in the shape ellipsoids. 
As we are dealing with oblate halos $b/a\approx 1$ and the
directions corresponding to the medium and major axis could be subject
to noise.
For this reason, in order to better quantify the twisting for each
halo we take the standard deviation of the angle between the two minor
axis, which should be less succeptible to noise.
With this measurement, twisting becomes a global property of the halo,
as the standard deviation is measured over the different radii at
which we have quantified the alignment.

Figure \ref{fig:alignment_correlations} shows the correlation between
twisting at the four different radii at which we have measured the
alignments. 
The median, mean vand standard deviation of the triaxiliy are $3$, $7$ and $8$
degrees, respectively.
The correlation between twisting and triaxility also changes with
radius as measured by the Spearman's rank correlation coefficient.
The largest correlation shows with the halo triaxility measured at
$0.12R_{200}$. 
The most important consequence of this changing degree in correlation
is that twisting cannot only be explained by large triaxiality values
and a possibly noisy determination of the shape axis.

@@

The presence of baryons produce rounder dark matter halos than its DMO
counterpart.
To better understand this relationship we quantify the correlation
between halo shape with baryonic disc properties.  
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disc to total mass ratio and
the optical radius.

Figure \ref{fig:disc_correlations} shows the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
We find that the strongest correlations are found with the halo shape
measured at radii smaller than $0.12R_{200}$.
The  trend is such that halos with large triaxiality correlate with
high gas density and stellar discs with low mass and small size. 
In turn, massive and large stellar discs within with low density of
gas in its core correlate with low halo triaxiliaty.

>>>>

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{array}

\usepackage{multirow}
\usepackage{multicol}
\usepackage{blindtext}
\newcolumntype{?}{!{\vrule width 1pt}}
\newcommand{\Msun}{\,{\rm M}$_{\odot}$\,}
\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
\newcommand{\kpch}{\,{\rm kpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
\newcommand{\kpc}{\,{\rm kpc}\,}

\title[Dark matter halo shapes at $z=0$] {Dark matter halo shapes at
  $z=0$ in the Auriga simulations} 
\author[Prada et al.]{
\parbox[t]{\textwidth}{
{Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}}
{Jaime E. Forero-Romero,$^{1}$}
{Robert J. J. Grand,$^{2}$}
{R\"udiger Pakmor,$^{2}$}
{Volker Springel$^{2}$}
}
\\\\
% List of institutions
$^{1}$Departamento de F\'isica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^{2}$Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzschild-Str. 1, D-85741 Garching, Germany\\
}

@@

% Abstract of the paper
\begin{abstract}
We present shape measurements of Milky Way sized dark matter halos at
redshift $z=0$ in a suite of 30 zoom simulations from the Auriga
project. 
We compare the results in full magnetohydrodynamics against dark
matter only simulations and find a strong influence of baryons 
in making dark matter haloes rounder at all radii compared to its dark 
matter only counterparts.
At distances $\lesssim 30$ kpc, rounder dark matter distributions
correlate with extended massive stellar discs and low core gas 
densities.  
We measure the alignment between the halo and the disc at
different radii and find a high degree of alignment at all radii for most 
of the galaxies.
In some cases the alignment significantly changes as a function of
radius implying that the halo shape twists; 
this effect correlates with extended stellar discs and
low gas densities and is almost absent in the dark matter only
simulations.  
In a comparison against observational constraints we find that $20\%$
of halos in our sample are consistent with observational results derived
from the Pal 5 stream that favor an almost spherical shape.
Including baryons is a required element to achieve this
level of agreement. In contrast, none of the simulations (dark matter
only nor with baryons) match the constrains derived from the
Sagittarius stream that favor an oblate dark matter halo.
\end{abstract}

@@

Our physical picture of the Universe as a whole has been shaped by
accurate observations and modeling of our own Galaxy. 
For instance, the study of the Milky Way (MW) morphology and its
separation into global kinematic coherent components, such as its
disc and bulge, has been used to support the existence of a Dark Matter (DM) component surrounding the Galaxy
\citep{2000MNRAS.311..361O,2009PASJ...61..227S,2010JCAP...08..004C,2013ApJ...779..115B,Iocco15}. 
In other words, explaining the matter budget and kinematic state of the MW also put constraints on the wider cosmological context that made possible its evolution.

A detailed study of the full three-dimensional MW
gravitational potential could constrain the properties of the DM halo surrounding our Galaxy and even help to pin down the nature of the DM particle.
For instance, the study of fossil records of stellar streams resulting from infalling globular clusters or
satellite galaxies that got tidally disrupted by the gravitational
potential of the MW, could be translated into tight constrains 
both on the shape of the dark matter halo in the outer regions of our Galaxy \citep{1998ApJ...495..297J,1999MNRAS.307..495H, 1999MNRAS.307..877T} and also bound the properties of dark matter substructures, which also depend on the microphysics of a given DM particle.

In any case, the observational constraints on the gravitational potential shape must be confronted against the expectations from different galaxy formation models in an explicit cosmological context.
For instance, in the current paradigm of a Cold Dark Matter
(CDM) dominated Universe galaxies are expected to be hosted by
triaxial DM halos. To what extent the CDM expectations are born out by observations in our Galaxy?

This question has been challenging to address because it has been difficult to produce realistic galactic discs in simulations within the CDM context \citep{1997ApJ...478...13N}.
The success of such enterprise has been based on understanding how baryonic effects, such as stellar feedback and black hole feedback, play an important role in forming a galactic disc resembling the MW.
Along the way, different numerical experiments have also shown that the baryonic
effects also impact the DM halo shape making it rounder
than it otherwise be in a DM-only simulation
\citep{Dubinski94,Debattista08,Kazantzidis10,Abadi10,Bryan13,Chua19,Artale19}. 
Another important aspect of the disc-halo relationship dynamics, namely, the radial evolution of the alignment between the stellar disc and the DM halo, has also been explored in some simulations \citep{Bailin05,Debattista13}.

@@

In this paper we analyze the internal DM halo shape in thirty MW type
galaxies from state-of-the-art simulations from the Auriga project.
These simulations have large enough numerical resolution, and explicit
cosmological context and an appropriate feedback physics to produce
realistic MW discs.
Furthermore, the relatively large number of simulated systems allows us
to have a handle on the statistical significance of our results.
This paper is structured as follows. 
In Section \ref{sec:numerical} we present the most relevant details of
the simulations, in Section \ref{sec:method} we present the method
that we use to measure the DM halo shape. 
In Section \ref{sec:results} we present our results focusing on the
radial shape trends at $z=0$ and the alignments with the stellar disc.
We discuss our results in Section \ref{sec:discussion} to place them
into the context of other numerical work, explore correlations of the
shape with baryonic properties in the disc and finally make a direct
comparison against observational constraints for the MW's dark matter halo shape.
We finalize with our conclusions in Section \ref{sec:conclusions}.

@@

The 30 zoom-in halos have a dark matter particle mass of $\sim 3\times
10^5$\Msun while the barynic mass resolution is $\sim 5\times 10^4$\Msun.
The softening lenght for gravitational force computation for stellar
particles and high-resolution dark matter particles 
is fixed to be 500 $h^{-1}$ pc in comoving coordinates up to $z=1$,
and 396 pc in physical coordinates afterwards.
The gravitational softening length for gas cells changes with the mean
cell radius but is limited to be larger than the stellar softening
lenght and 500 $h^{-1}$ pc comoving. 
From now on we refer to the haloes simulated with baryonic physics as the
MHD sample and to the haloes simulated with dark matter only as the
DMO sample.

\begin{figure*}
\includegraphics[width=0.45\textwidth]{level4_DM_halo_24_2.png}
\includegraphics[width=0.45\textwidth]{level4_DM_halo_24_4.png}
\includegraphics[width=0.45\textwidth]{level4_MHD_halo_24_2.png}
\includegraphics[width=0.45\textwidth]{level4_MHD_halo_24_4.png}
  \caption{DM density in logarithmic scale within a slice of one tenth
    of the virial radius.
    The cut is perpendicular to the short axis of the inertia tensor ellipsoid.
    The black ellipses show the results of the fitting procedure
    described in Section \ref{sec:method}. 
    Upper/lower panels correspond to DMO/MHD simulations, respectively.
    Left/right panels show data at small/large radii, respectively.
    This plot showcases the most noticeable effect in all halos
    across the Auriga simulations: haloes are rounder at all radii
    after baryonic physics is included.}
\label{fig:slices}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}
\end{center}
\caption{\textbf{RP: Maybe add DMO and MHD directly to the two panels?} Axial ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $14\pm 1$ kpc
  ($230\pm 15$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder than its DMO counterparts.
  Third, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend). \textbf{RP: is that last point really statistically the case? The average points that were in the plot before seemed to be consistent within scatter for the MHD run} }
  \label{fig:triaxiality_plane}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_DM.pdf}
\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_MHD.pdf}
\caption{Triaxility cumulative distribution at five different radii.
  Right/left panel corresponds to DMO/MHD simulations, respectively. 
  In DMO simulations the median triaxiality at all radii is larger
  than $2/3$. Furthermore, the triaxility increases as one moves
  towards the inner part of the halo.
  In MHD simulations these trend reverses.
  The median triaxility at all radii is smaller than $1/3$ and the
  halo is less triaxial as on moves towards the stellar disc.}
\label{fig:triaxial_cumulative}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{angles_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{\textbf{RP: I think the upper row that compares the shape of the halo from the DMO sim with the orientation of the disk in the MHD sim is hard to interpret. Maybe instead show in the first row the angle against a fixed axis (e.g. total angular momentum of the halo?)} Angles between the principal axis of the dark matter halo
  shape and the angular momentum of the stellar disc 
  as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major/middle/minor axis in the halo.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value at every radius.
  In the upper row the haloes come from the DMO simulation, 
  showing that the directions in the ellipsoids describing the shape
  are constant as a function of radius for the most part of the sample.
  In the lower row the haloes come from the MHD simulation providing a 
  self-consistent comparison with the stellar discs. 
  In this case the dark matter shells twists significantly in $1/5$ of the sample. For the remaining haloes the alignment is almost constant as a function of radius.
}
\label{fig:cumulative_alignment}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{correlations_twisting_triaxiality_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alignment (measured as the standard deviation of the alignment angles)
  as a function of the halo triaxialiy at the radius indicated
  in each panel's title.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient while the line shows the result of the
  minimum squares fit to the data.
  Twisting does not correlate with triaxility in the same way at every
  radii. The strongest correlation appears around $0.12R_{200}$ ($28\pm2$ kpc).} 
\label{fig:alignment_correlations}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the same
  baryonic disc properties in Figure \ref{fig:disc_correlations}.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best minimum squares fit to a line.
  In this case the disc to total mass ratio does not correlate with the alignment twisting
  The gas density at the center and the disc optical radius still show
  a strong correlation, albeit weaker than what they had with the triaxility.}
\label{fig:alignment_correlations2}
\end{figure*}

@@

The DM halo shape is an estimate of either the isopotential or
isodensity surfaces.   
Observational inference models usually estimate the 
isopotential contours which are probed by tracers (gas, stars), while
simulations work with the isodensity contours which can be directly
calculated from particle positions.  
Furthermore, the density contours in thin shells are very sensitive to
the presence of small satellites.  
For this reason we measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.  
This method yields results in good agreement to the isodensity
contours for radii $\leq 140$ kpc as explored by
\citep{VeraCiro11}.

@@

Following the convergence criterion by \cite{VeraCiro11} in DM only simulations
we restrict the measurement of the ellipsoidal parameters to radii
between $2$ kpc and $R_{200}$ where  $R_{200}$ corresponds to the radius
computed enclosing a sphere with 200 times the critical density of the
Universe. 
This scale is similar to the gravitational smoothing length for
stellar particles. 
For this reason, in order to make the DM halo shape results comparable
against stellar disc shape measurements, we report our DM halo shape
results above $10$ kpc.
We characterize the stellar disk by measuring the angular momentum vector using the $10\%$ oldest stellar particles.

@@

We summarize this trend in Figure \ref{fig:triaxiality_plane} by
plotting the results of all the 30 halos in the DMO sample.
The left panel shows every halo in the $c/a$-$b/a$ plane at
two different radii $R_{200}/8 (\sim 20$kpc$)$ \textbf{RP: the plots says R200/16} and $R_{200}$. 
The outer part of the halo is systematically rounder than its inner
region. 
Nevertheless the halo shape can still be considerated to be prolate at
all radii.

@@

The left panel of Figure \ref{fig:triaxial_cumulative} shows that in
the DMO sample the triaxility has a median larger than $2/3$ (a
typical value that marks the transition from triaxiality to
prolateness) at all radii. 
Furthermore this median value increases as we move towards the inner part of the
halo.
The right panel shows how the correlations in the MHD simulations go
in the opposite direction with respect to DMO result.s
There the median triaxility is always smaller than $1/3$ (a typical
values that marks the transition from triaxility to prolateness) and this
median value decreases as we move towards the inner part of the halo.

We now quantify this change in triaxility at the level of individual halos.
We compute $\Delta T\equiv T_{\rm MHD}-T_{\rm DMO}$ to quantify the
change between the triaxility in the MHD and the DMO simulation.
Figure \ref{fig:delta_triaxial_cumulative} shows the cumulative
distribution at the same radii as in
Figure \ref{fig:delta_triaxial_cumulative}.  
From this Figure we observe that at $R_{200}$ all the halos have
reduced its triaxility include baryonic physics.
At smaller radii this general trend continues, although there is a
small fraction ($\approx 1/15$) \textbf{RP: just write 2 in out sample?} of halos that slightly increased their
triaxility with MHD physics. 
However, this small fraction correspond to halos that already were
outliers in the DMO sample and had triaxilities around $1/3$,
considerably lower than its parent sample.

@@

A common assumption in dynamical models of the MW DM halo is that
its angular momentum is perfectly aligned with the stellar disc minor axis.
Although it is a reasonable assumption to guarantee the stability of
the galactic disc in simplified models of isolated galaxies, this
might not perfectly hold in an explicit cosmological context. 
To examine the degree of validity of this assumption we study in this
subsection the alignment between the stellar disc's angular momentum
eigenvectors of the dark matter halo shape.

The upper row in Figure \ref{fig:cumulative_alignment} shows 
the alignments between the major/intermediate/minor axis of the
halo and the disc angular momentum as a function of the dimensionless radius at which the halo shape was measured.
The upper row shows the angle between the DMO halo and the stellar
disk in the MHD simulation.
This comparison help us to measure how the shape directions in the
dark matter shells change with radius.
For the bulk of the sample the lines are horizontal.
Only two halos show a significant change of more than 30 degrees across different radii.

The lower row in Figure \ref{fig:cumulative_alignment} shows the
results for the MHD simulations.
The most striking feature of this plot is that most of the discs have their angular momentum aligned with the minor axis of the halo and perpendicular to the intermediate and major axis.
The median and the standard deviation of the alignment angles with
the major/median/minor axis are $82\pm14$, $82\pm16$ and $19\pm20$
degrees, respectively.

In six halos the alignment changing significantly as a function of radius, this can be seen as a twisting in the shape ellipsoids. 
As we are dealing with oblate halos $b/a\approx 1$ and the
directions corresponding to the medium and major axis could be subject
to noise.
For this reason, in order to better quantify the twisting for each
halo we take the standard deviation of the angle with the minor
axis, which should be less susceptible to noise.
With this measurement, twisting becomes a global property of the halo,
as the standard deviation is measured over the different radii at
which we have quantified the alignment.

Figure \ref{fig:alignment_correlations} shows the correlation between
twisting at the four different radii at which we have measured the
alignments. 
The median, mean and standard deviation of the twisting are $2$, $6$ and $8$ degrees, respectively.
The correlation between twisting and triaxility also changes with
radius as measured by the Spearman's rank correlation coefficient.
The largest correlation shows with the halo triaxility measured at
$0.12R_{200}$. 
The most important consequence of this changing degree in correlation
is that twisting cannot only be explained by large triaxiality values
and a possibly noisy determination of the shape axis.

@@

The presence of baryons produces rounder dark matter halos than its DMO
counterpart.
To better understand this relationship we quantify the correlation
between halo shape with baryonic disc properties.  
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disc to total mass ratio and
the optical radius.

Figure \ref{fig:disc_correlations} shows the correlations of
these quantities with the triaxility at five different radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
We find that the strongest correlations are found with the halo shape
measured at radii smaller than $0.12R_{200}$.
The  trend is such that halos with large triaxiality correlate with
high gas density and stellar discs with low mass and small size. 
In turn, massive and large stellar discs within with low density of
gas in its core correlate with low halo triaxiliaty.
